---
title: "Group4"
author: "Emmanuel Kossy Okonkwo Vicky Caravaggio James Tuttle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Group4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette demonstrates how to use the functions provided in the Group4 package to:

- Prepare the forest fire dataset.
- Estimate logistic regression coefficients using a custom optimization routine (without relying on built-in logistic regression functions).
- Compute bootstrap confidence intervals for the estimated coefficients.
- Construct a confusion matrix and derive performance metrics from the fitted logistic model.


# Data and Package Setup 

The forest fires dataset is included in the Group4 package and can be accessed via system.file. The dataset includes various weather and geographic variables, and the binary response variable binary_area indicates whether the burned area is greater than zero.

```{r setup}
library(Group4)
```

# Data Preparation

We begin by preparing the forest fires data. The prepare_forestfires_data function reads the CSV file included in the package and processes it by creating a binary response variable (binary_area), and dropping unnecessary columns:

```{r message=FALSE, warning=FALSE}
file_path <- "forestfires.csv"
forestfires_data <- prepare_forestfires_data(file_path)
head(forestfires_data)

```

The resulting dataframe contains numerical predictors and a binary response (0/1).


# Estimating Coefficients

We estimate logistic regression coefficients using the estimate_beta function. This function:

- Prepares the data (converts factors to numeric, handles date variables).
- Uses the least squares solution as starting values.
- Optimizes the negative log-likelihood of a logistic regression model using optim.


```{r}
result <- estimate_beta(forestfires_data, "binary_area")
result$beta

```

The vector result$beta provides the estimated coefficients.


# Bootstrap Confidence Intervals

We can compute bootstrap confidence intervals for these coefficients using the bootstrap_ci function. By default, it performs 20 bootstrap resamples and provides a 95% confidence interval (you can adjust these parameters).

```{r}
ci <- bootstrap_ci(forestfires_data, "binary_area", result$beta, n_bootstrap = 20, alpha = 0.05)
ci

```
The result ci shows the lower and upper percentile-based confidence intervals for each estimated coefficient.



# Confusion Matrix and Performance Metrics

Once we have a fitted model, we can use confusion_matrix to generate a confusion matrix at a cut-off of 0.5. This function returns:

- A 2x2 confusion matrix (actual vs. predicted).
- Prevalence
- Accuracy
- Sensitivity
- Specificity
- False Discovery Rate
- Diagnostic Odds Ratio


```{r}
metrics <- confusion_matrix(forestfires_data, "binary_area", result$beta)
metrics$confusion_matrix

```
We can also print out the derived metrics:

```{r}
cat("Prevalence:", metrics$prevalence, "\n")
cat("Accuracy:", metrics$accuracy, "\n")
cat("Sensitivity:", metrics$sensitivity, "\n")
cat("Specificity:", metrics$specificity, "\n")
cat("False Discovery Rate:", metrics$false_discovery_rate, "\n")
cat("Diagnostic Odds Ratio:", metrics$diagnostic_odds_ratio, "\n")

```


# Summary
In this vignette, we demonstrated a complete workflow:

- Time-series data set preparation
- Logistic regression coefficient estimation (custom optimization)
- Bootstrap-based confidence intervals
- Confusion matrix and relevant metrics

This workflow provides a foundation for experimenting with custom logistic regression implementations and gaining more transparent insight into model performance.



# References

You can find the references [here](https://chatgpt.com/share/675202cb-5bac-8010-b516-8dcecc495f9a)
Before installing the 'Group4' package, ensure the 'devtools' and 'repo' libraries are active. The package can be retrieved from the GitHub repository found [here](https://github.com/AU-R-Programming/FinalProject_Group4).
