---
title: "Group4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

This vignette demonstrates how to use the `Group4` package for supervised binary classification using logistic regression.

# Loading the Package

```{r}
library(Group4)
```

# Example Usage with Simulated Data


Generating Simulated Data

```{r}
set.seed(123)
data <- data.frame(
  x1 = rnorm(100),
  x2 = as.factor(sample(c("A", "B"), 100, replace = TRUE)),
  x3 = as.Date('2023-01-01') + sample(1:100, 100),
  y = rbinom(100, 1, 0.5)
)

```

Estimating Beta Coefficients

```{r}
result <- estimate_beta(data, "y")
cat("Estimated Beta Coefficients:\n")
print(result$beta)

```

Calculating Bootstrap Confidence Intervals

```{r}
ci <- bootstrap_ci(data, "y", result$beta, n_bootstrap = 100, alpha = 0.05)
cat("\nBootstrap Confidence Intervals:\n")
print(ci)

```

Generating Confusion Matrix and Metrics

```{r}
metrics <- confusion_matrix(data, "y", result$beta)
cat("\nConfusion Matrix:\n")
print(metrics$confusion_matrix)
cat("\nMetrics:\n")
cat("Prevalence:", metrics$prevalence, "\n")
cat("Accuracy:", metrics$accuracy, "\n")
cat("Sensitivity:", metrics$sensitivity, "\n")
cat("Specificity:", metrics$specificity, "\n")
cat("False Discovery Rate:", metrics$false_discovery_rate, "\n")
cat("Diagnostic Odds Ratio:", metrics$diagnostic_odds_ratio, "\n")

```


# Conclusion

In this vignette, we demonstrated how to use the Group4 package to perform logistic regression analysis, calculate bootstrap confidence intervals, and evaluate model performance using a confusion matrix and associated metrics.


# References

you can find the reference [here](https://chatgpt.com/share/674f9710-950c-8010-ab01-c910f66608bb)


